---
title: 'Part2:Part 2: Basic_Inferential_Data_ Analysis'
author: "Aleksander Petrovskii"
date: "June 11, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls(all=TRUE))
library(ggplot2)
```
## Overview
***
The purpose of the this data analysis is to analyze the ToothGrowth data set by comparing the guinea tooth growth by supplement and dose. First, I will do exploratory data analysis on the data set. Then I will do the comparison with confidence intervals in order to make conclusions about the tooth growth.

## Exploring the Data (Question 1 and 2)
***  
#### Load data
```{r echo=TRUE}

library(datasets)
data(ToothGrowth)
str(ToothGrowth)
```
#### Look at data

```{r echo=TRUE}
head(ToothGrowth)

```

#### Provide basic summary of dataset

```{r echo=TRUE}
summary(ToothGrowth)
```
#### Combining the dosage and the delivery method to analyse the statistical data

```{r echo=TRUE}
    by(ToothGrowth$len, INDICES = list(ToothGrowth$supp, ToothGrowth$dose), summary)

```


#### Look at possible relation between tooth length and delivery methods

```{r echo=TRUE}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
  geom_boxplot() +
  facet_grid(. ~ supp) +
  xlab("Dosage (miligrams)") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Supplement type"))

```
  
####  Look at possible relation between tooth length vs the delivery methods  by dose amount
   
```{r echo=TRUE}
ggplot(ToothGrowth, aes(x=supp, y=len)) +
        geom_boxplot(aes(fill=factor(supp)))+
        geom_point() +
        facet_grid(.~dose) +
        xlab("Supplement Delivery") + 
        ylab("Tooth Length") + 
        theme(plot.title = element_text(hjust = 0.5)) +
        ggtitle("Tooth Length by Supp & Dosage")

```



## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose (Question 3)
***    
#### First, let`s analyze Dosage as a factor

```{r echo=TRUE}
    dosage_1 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
    dosage_2 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
    dosage_3 <- subset(ToothGrowth, dose %in% c(0.5, 2.0))
    ```
#### For dosage from 0.5 to 1.0 (milligrams)
```{r echo=TRUE}
t.test(len ~ dose, paired = F, var.equal = F, data = dosage_1)

```
#### For dosage from 1.0 to 2.0 (milligrams)
```{r echo=TRUE}
t.test(len ~ dose, paired = F, var.equal = F, data = dosage_2)

```
#### For dosage from 0.5 to 2.0 (milligrams)
```{r echo=TRUE}
t.test(len ~ dose, paired = F, var.equal = F, data = dosage_3)

```

### Second, let`s analyze Supplement as a factor
***   
```{r echo=TRUE}
t.test(len ~ supp, paired = F, var.equal = F, data = ToothGrowth)
```
## Conclusions and assumptions (Question 3)  
***   
#### Assumptions
1. For Dosage as a factor the confidence intervals ([-11.98, -6.276] for doses 0.5 and 1.0, [-18.16, -12.83], and [-8.996, -3.734] for doses 1.0 and 2.0) and for doses 0.5 and 2.0,  allow for the rejection of the null hypothesis and a confirmation that there is a significant correlation between tooth length and dose levels.
2. For Supplement as a factor the confidence interval of [-0.171, 7.571] does not allow us to reject the null hypothesis (that there is no correlation between delivery method and tooth length).


#### Conclusions
1. Increasing the dose level leads to increased tooth growth
2. Supplement type has no effect on tooth growth.





